---
title: "`r params$title`"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  html_document:
    toc: true
    toc_float: true
params:
  runs_file: NULL
  title: "cortisol.nmea simulation report"
  full_method: "Full"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 3.6)
library(cortisol.nmea)
runs <- readRDS(params$runs_file)
d <- runs$design
fmt <- function(x, digits = 3) formatC(x, digits = digits, format = "f")
has_full <- params$full_method %in% names(d$methods) &&
  !is.null(runs$cases[[1]]$methods[[params$full_method]]$points)
```

## Design

```{r design}
cat_rows <- data.frame(
  Item = c("Trimesters", "Subjects per cohort", "Base cohorts per trimester", "Scenarios", "Methods",
           "Evaluation grid"),
  Value = c(paste(d$trimesters, collapse = ", "), d$n, paste(lengths(d$seeds), collapse = " / "),
            paste(names(d$scenarios), collapse = ", "), paste(names(d$methods), collapse = ", "),
            sprintf("%s to %s h, %d points", min(d$times), max(d$times), length(d$times))))
knitr::kable(cat_rows)
```

Each seed defines one base cohort per trimester; the scenarios of a seed share its
subjects, sampling times and clean noise. The GAMM comparator is an adaptation of
Sánchez et al. (2012), not the original authors' code.

```{r failures}
fail <- do.call(rbind, lapply(runs$cases, function(cs) {
  do.call(rbind, lapply(names(cs$methods), function(m) {
    r <- cs$methods[[m]]
    if (r$status == "ok") NULL else data.frame(Case = cs$id, Method = m, Error = r$error)
  }))
}))
if (is.null(fail)) {
  cat("All method fits completed.\n")
} else {
  knitr::kable(fail, caption = "Failed fits (kept as failures, never replaced)")
}
```

## Curve recovery

Within each cohort, errors are averaged over the subjects available for all methods,
then averaged over cohorts. Intervals are 95% Monte Carlo t intervals over cohorts;
the "All" row weights trimesters equally. RMSE is in nmol/L; the curve-AUC error is
the absolute error of the 0-18 h curve integral in nmol*h/L (not ROC-AUC).

```{r recovery-table}
s <- eval_summary(runs)$summary
s <- s[s$measure %in% c("rmse", "auc_error"), ]
s$estimate <- sprintf("%s [%s, %s]", fmt(s$mean), fmt(s$lower), fmt(s$upper))
tab <- reshape(s[, c("trimester", "scenario", "method", "measure", "valid", "planned", "estimate")],
               idvar = c("trimester", "scenario", "method", "valid", "planned"), timevar = "measure",
               direction = "wide")
tab <- tab[order(match(tab$trimester, c("All", d$trimesters)), tab$scenario, tab$method), ]
knitr::kable(tab, row.names = FALSE, col.names = c("Trimester", "Scenario", "Method", "Valid cohorts",
                                                    "Planned", "RMSE", "Curve-AUC error"))
```

```{r recovery-plots, fig.height = 3.4}
print(plot_recovery(runs, "rmse"))
print(plot_recovery(runs, "auc_error"))
```

```{r paired, eval = has_full}
others <- setdiff(names(d$methods), params$full_method)
pr <- do.call(rbind, lapply(others, function(o) eval_paired(runs, params$full_method, o)$summary))
pr <- pr[pr$trimester == "All", ]
pr$estimate <- sprintf("%s [%s, %s]", fmt(pr$mean), fmt(pr$lower), fmt(pr$upper))
knitr::kable(pr[, c("scenario", "method", "valid", "estimate", "wins")], row.names = FALSE,
             caption = "Paired RMSE differences (pooled over trimesters; negative favours the first method)",
             col.names = c("Scenario", "Comparison", "Valid cohorts", "Mean difference [95% MC]", "Wins"))
```

## Example cohort

```{r cohort-curves, fig.height = 3.2}
print(plot_cohort_curves(runs))
```

## Outlier detection and subject retention

```{r detection, eval = has_full}
det <- eval_detection(runs, params$full_method)$summary
det <- det[det$measure %in% c("FPR", "FNR", "precision", "roc_auc", "cutoff"), ]
det$value <- sprintf("%s [%s, %s]", fmt(det$mean), fmt(det$lower), fmt(det$upper))
knitr::kable(reshape(det[, c("trimester", "scenario", "measure", "value")],
                     idvar = c("trimester", "scenario"), timevar = "measure", direction = "wide"),
             row.names = FALSE, caption = "Detection by the initial-residual outlier step")
```

```{r detection-plots, eval = has_full, fig.width = 4.5, fig.height = 3.4}
contaminated <- setdiff(names(d$scenarios), "D0")
for (sc in contaminated) {
  print(plot_detection(runs, params$full_method, sc))
  print(plot_retention(runs, params$full_method, sc))
}
```

```{r retention-table, eval = has_full}
ret <- eval_retention(runs, params$full_method)$summary
ret$value <- fmt(ret$mean, 1)
knitr::kable(reshape(ret[, c("trimester", "scenario", "measure", "value")],
                     idvar = c("trimester", "scenario"), timevar = "measure", direction = "wide"),
             row.names = FALSE, caption = "Subjects by workflow outcome (%, mean over cohorts)")
```

Recovery is evaluated on retained subjects; excluded subjects are not recovered by the
workflow and are reported here so that lower errors are read together with coverage.

## Session

```{r session}
cat(sprintf("cortisol.nmea %s, saemix %s, mgcv %s, %s\n", utils::packageVersion("cortisol.nmea"),
            utils::packageVersion("saemix"), utils::packageVersion("mgcv"), R.version.string))
```
