---
title: "Benchmark Metrics Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmark Metrics Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cortisol.nmea)
```

Simulation benchmarks compare estimated curves against a known synthetic
truth. This vignette uses a tiny example to demonstrate the high-level method
comparison workflow and the metric table helpers.

```{r simulation}
study <- run_simulation_study(
  n = 20,
  alpha = 1.25,
  missing_scenarios = c("S0_CleanObserved", "S4_SparseMorningMissing"),
  contamination_scenarios = "SC2_TimingError",
  seed = 201
)
```

Fast truth-based comparison methods are always available:

- `OracleTruth`: evaluates the true synthetic parameters.
- `SymmetricTruth`: evaluates the true parameters after setting `alpha = 1`.

The optional `GAM` baseline is included when `mgcv` is installed.

```{r compare}
methods <- c("OracleTruth", "SymmetricTruth")
if (requireNamespace("mgcv", quietly = TRUE)) {
  methods <- c(methods, "GAM")
}

comparison <- run_method_comparison(
  study,
  methods = methods,
  scenarios = c("S0_CleanObserved", "S4_SparseMorningMissing", "SC2_TimingError")
)

comparison$summary
```

Subject-level metrics are stored separately from method-level summaries.

```{r subject-metrics}
head(comparison$subject_metrics)
```

`summarize_winners()` gives a compact winner table for one or more metrics.

```{r winners}
summarize_winners(comparison$summary)
```

`make_benchmark_tables()` returns a long table, a wide table, and the winner
summary together.

```{r tables}
tables <- make_benchmark_tables(comparison$summary)
names(tables)
head(tables$long)
```

The fitted NMEA methods, `Baseline_NMEA` and `NMEA_AlphaOnly`, are available
when `saemix` is installed. They are deliberately not run in this vignette so
the package examples stay fast.
