---
title: "Benchmark Metrics Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmark Metrics Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cortisol.nmea)
```

Simulation benchmarks compare estimated curves against a known synthetic
truth. This vignette uses a tiny example to demonstrate the metric layer.

```{r truth}
time_grid <- seq(0, 18, by = 0.25)
psi_true <- simulate_nmea_parameters(n = 25, alpha = 1.25, seed = 201)
truth <- model_eval(time_grid, psi_true)
```

We construct two simple candidate methods without running a full model fit:

- `AlphaOnly`: the true parameters except for a misspecified alpha.
- `NoisyOracle`: the true curves with mild independent prediction noise.

```{r candidates}
psi_alpha_only <- psi_true
psi_alpha_only[, "alpha"] <- 1
pred_alpha_only <- model_eval(time_grid, psi_alpha_only)

set.seed(202)
pred_noisy_oracle <- truth + matrix(
  rnorm(length(truth), mean = 0, sd = 0.35),
  nrow = nrow(truth),
  dimnames = dimnames(truth)
)
```

Subject-level metrics are produced with `curve_benchmark_metrics()`.

```{r subject-metrics}
m_alpha <- curve_benchmark_metrics(pred_alpha_only, truth, time_grid)
m_oracle <- curve_benchmark_metrics(pred_noisy_oracle, truth, time_grid)

head(m_alpha)
```

Aggregate metrics are produced with `summarize_curve_metrics()`.

```{r aggregate}
rbind(
  cbind(Method = "AlphaOnly", summarize_curve_metrics(m_alpha)),
  cbind(Method = "NoisyOracle", summarize_curve_metrics(m_oracle))
)
```

These helpers are deliberately small. Larger benchmarking workflows can build
on top of them by adding method labels, scenario labels, convergence metadata,
coverage penalties, and tie-aware winner summaries.
