---
title: "Fitting diurnal cortisol curves"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting diurnal cortisol curves}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 3.5)
```

```{r setup}
library(cortisol.nmea)
```

## The model

Each person's cortisol curve is a scaled logistic bump on a baseline,

$$f(t) = c_0 + \frac{c_1}{s}\, g\!\left(\frac{t^* - \mu}{s}\right),$$

where $g$ is the logistic density and $t$ is hours since waking. $\mu$ is the
time of the peak, $s$ its width, $c_1$ its size and $c_0$ the baseline. The rise
before the peak is often steeper than the decline after it, so time before
$\mu$ is rescaled, $t^* = \alpha t + (1-\alpha)\mu$, with $\alpha > 1$ giving a
sharper rise. The four parameters vary between people (random effects) and
$\alpha$ is shared within a fit.

```{r curve}
psi <- rbind(
  symmetric = c(mu = 0.6, s = 0.75, c1 = 30, c0 = 2.5, alpha = 1),
  steeper_rise = c(mu = 0.6, s = 0.75, c1 = 30, c0 = 2.5, alpha = 1.5)
)
t <- seq(0, 18, by = 0.05)
matplot(t, t(nmea_curve(t, psi)), type = "l", lty = 1, col = c("grey50", "#277DA8"),
        xlab = "Hours since waking", ylab = "Cortisol (nmol/L)")
legend("topright", c("alpha = 1", "alpha = 1.5"), lty = 1, col = c("grey50", "#277DA8"), bty = "n")
```

## Preparing data

The data are in long format, one row per sample. `cortisol_data()` checks them
and standardizes the column names. Time is hours since the first sample of the
day, which is taken as waking; concentrations are in nmol/L.

Here a simulated second-trimester cohort of 40 people stands in for real data.
It contains some aberrant samples, as real data often do.

```{r data}
sim <- sim_contaminate(sim_cohort(sim_reference("T2"), n = 40, seed = 2026), "D2")
d <- sim$observed
d
```

With your own data, the call looks like

```r
d <- cortisol_data(my_data, id = "participant", time = "hours_since_waking",
                   value = "cortisol", trimester = "trimester")
```

`convert_cortisol_units()` converts from ug/dL or ng/mL, and `prep_cortisol()`
can restrict the data to a time window. Fit each trimester or study wave
separately.

## Running the workflow

`nmea_pipeline()` runs the full workflow by default:

1. fit the model with $\alpha = 1$;
2. remove samples whose absolute residual exceeds a cutoff;
3. exclude people left with fewer than three samples;
4. refit, and exclude people whose curve explains less than half of their
   variation (fraction of variance unexplained, FVU, of 0.5 or more) or whose
   curve has no morning peak ($c_1 \le 0$);
5. refit over $\alpha = 0.8, 0.9, \dots, 1.5$ and keep the best fit.

```{r fit}
fit <- nmea_pipeline(d)
fit
```

The cutoff in step 2 is $k$ times a noise standard deviation. For data labelled
T1, T2 or T3 the package uses 3.5 times the noise SD of that trimester's
reference; otherwise it estimates the noise SD from the data. See
`vignette("thresholds")`.

The result records what happened to every sample and every person.

```{r records}
head(fit$points[fit$points$flagged, ])
table(fit$subjects$outcome)
```

People who were excluded keep their rows, with no curve. Report how many were
excluded and why alongside any result based on the fitted curves.

## Curves and features

```{r predict}
curves <- predict(fit, times = seq(0, 18, by = 0.1))
dim(curves)
```

```{r features}
features <- nmea_features(fit)
head(round(features, 2))
```

| Feature | Meaning |
|---|---|
| `AUC` | area under the curve from waking to 18 h (nmol·h/L) |
| `EML` | level at waking |
| `PCL` | peak level |
| `AR` | awakening rise, `PCL - EML`, or 0 if the peak is before waking |
| `DDC` | steepest decline after the peak (nmol/L per hour) |

## Changing the workflow

Each step is controlled by `nmea_steps()`; set a step to `NULL` to skip it.

```{r steps}
nmea_steps(outlier = outlier_fixed(6))              # the original fixed cutoff
nmea_steps(outlier = outlier_sd(2.75, "mad"), fvu_max = NULL)
nmea_steps(alpha_grid = 1)                          # symmetric curves only
```

`nmea_steps_direct()` fits all samples with no removal or screening, which is
useful as a comparison.

```{r direct}
direct <- nmea_pipeline(d, steps = nmea_steps_direct())
direct
```

Fitting settings (SAEM iterations, seed, starting values) are in
`nmea_control()`. The fit uses the `saemix` package; results are reproducible
for a given seed, and your own random number stream is left unchanged.

## A simulated cohort lets us check the result

Because this cohort was simulated, the true curves are known.

```{r check}
truth <- nmea_curve(seq(0, 18, by = 0.1), sim$truth$psi)
rmse <- function(pred) sqrt(rowMeans((pred - truth)^2))
kept <- fit$subjects$outcome == "retained"
c(full = mean(rmse(curves)[kept]),
  direct = mean(rmse(predict(direct, times = seq(0, 18, by = 0.1)))[kept]))
```

`vignette("simulation")` shows how to do this systematically.
