---
title: "The NMEA Cortisol Curve Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The NMEA Cortisol Curve Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
```

```{r setup}
library(cortisol.nmea)
```

The NMEA model represents a diurnal cortisol curve with a scaled logistic
shape and an optional pre-peak time transformation. The parameter vector is:

- `mu`: peak-location control.
- `s`: width or steepness control.
- `c1`: curve magnitude.
- `c0`: baseline level.
- `alpha`: pre-peak asymmetry control.

The symmetric curve is evaluated with `sl_model()`. The NMEA asymmetric curve
is evaluated with `asym_sl_model()`.

```{r curve-eval}
psi_one <- c(mu = 1.0, s = 1.1, c1 = 32, c0 = 2, alpha = 1.25)
time_grid <- seq(0, 18, by = 0.1)

yhat <- asym_sl_model(time_grid, psi_one)
head(yhat)
```

```{r curve-plot}
plot(
  time_grid,
  yhat,
  type = "l",
  xlab = "Hours since waking",
  ylab = "Cortisol",
  main = "Synthetic NMEA cortisol curve"
)
```

For multiple subjects, store one parameter vector per row and use
`model_eval()`.

```{r multi-subject}
psi <- simulate_nmea_parameters(n = 5, seed = 1)
curves <- model_eval(time_grid, psi)
dim(curves)
```

The package also provides derived features used by the project workflow.

```{r features}
summary_features(psi)
```

Model fitting is available with `fit_nmea()` when the optional `saemix`
dependency is installed. The example below uses only four synthetic subjects
and short SAEM settings so the vignette stays lightweight.

```{r fit-example, eval = requireNamespace("saemix", quietly = TRUE)}
data(toy_cortisol)
toy_small <- subset(toy_cortisol, Subject %in% unique(Subject)[1:4])

fit <- fit_nmea(
  toy_small,
  id = "Subject",
  time = "Time",
  response = "Conc_Obs",
  alpha = 1.2,
  control = list(
    nbiter.saemix = c(20, 10),
    nbiter.burn = 10,
    nbiter.map = 10,
    fim = FALSE
  )
)

summary_features(fit$psi)
```
