---
title: "Reproducing the paper's simulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducing the paper's simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

This vignette lists the exact design of the simulation study in the paper. The
code is not run when the vignette is built; the full study takes about two
hours on three cores.

```{r setup}
library(cortisol.nmea)
```

## Design

Each trimester has ten independent cohorts of 100 people. Every cohort is
analysed without contamination (D0), with mild contamination (D1: 5% of
samples raised by 4-8 noise SDs) and with strong contamination (D2: 10% raised
by 8-12 noise SDs).

```{r design}
ctrl <- nmea_control()   # 300 + 100 SAEM iterations, seed 632545
methods <- list(
  Direct = method_nmea(nmea_steps_direct(), ctrl),
  Full = method_nmea(nmea_steps(), ctrl),                               # default rule
  Full_fixed6 = method_nmea(nmea_steps(outlier = outlier_fixed(6)), ctrl),
  Full_MAD = method_nmea(nmea_steps(outlier = outlier_sd(2.75, "mad")), ctrl),
  GAMM = method_gamm()
)
design <- sim_design(
  trimesters = c("T1", "T2", "T3"), n = 100,
  seeds = list(T1 = 610161001:610161010, T2 = 610162001:610162010, T3 = 610163001:610163010),
  scenarios = c("D0", "D1", "D2"), methods = methods
)
```

## Running

```{r run}
future::plan(future::multisession, workers = 3)
runs <- sim_run(design, cache_dir = "paper-simulation-cache", parallel = TRUE)
saveRDS(runs, "paper-simulation-runs.rds")
```

## Tables and figures

Curve recovery on the people all compared methods could fit:

```{r tables}
main <- c("Direct", "Full", "GAMM")
eval_summary(runs, methods = main)$summary
eval_paired(runs, "Full", "Direct")$summary
eval_paired(runs, "Full", "GAMM")$summary
eval_paired(runs, "Full", "Full_fixed6")$summary
eval_retention(runs, "Full")$summary
eval_detection(runs, "Full")$summary
```

Figure 1 shows one cohort under the three scenarios and individual examples;
Figure 2 shows recovery across all cohorts, detection errors and retention.

```{r figures}
plot_cohort_curves(runs, "T3", seed = 610163001, methods = main)
plot_subject(runs, "T3_S0013", trimester = "T3", seed = 610163001, scenario = "D2", methods = main)
plot_recovery(runs, "rmse", methods = main)
plot_recovery(runs, "auc_error", methods = main)
plot_detection(runs, "Full", "D2")
plot_retention(runs, "Full", "D2")
sim_report(runs, "paper-simulation-report.html")
```

The fixed-cutoff and MAD arms are reference comparisons for the choice of the
outlier rule.
