---
title: "Synthetic Simulation Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Synthetic Simulation Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
```

```{r setup}
library(cortisol.nmea)
```

This vignette shows the public, synthetic-only simulation workflow. It mirrors
the structure of the Simulation V3 research scripts without depending on real
participant data. For most package users, `run_simulation_study()` is the
entry point: it creates latent curves, observed measurements, missing/sparse
scenarios, and contamination scenarios in one reproducible object.

## Step 1: Run a study

```{r study}
study <- run_simulation_study(
  n = 40,
  alpha = 1.2,
  missing_scenarios = c("S0_CleanObserved", "S4_SparseMorningMissing"),
  contamination_scenarios = c("C1_Spike", "SC2_TimingError"),
  seed = 101
)

names(study$scenarios)
```

The returned object keeps the synthetic subject parameters, latent truth,
observed clean data, scenario data sets, and configuration metadata together.

```{r object-shape}
str(study$config)
dim(study$latent$curve_mat)
head(study$observed)
```

## Step 2: Inspect latent curves

```{r latent-plot}
matplot(
  study$latent$time_grid,
  t(study$latent$curve_mat[1:10, ]),
  type = "l",
  lty = 1,
  col = "#1f6fb255",
  xlab = "Hours since waking",
  ylab = "Cortisol",
  main = "First 10 synthetic latent curves"
)
```

## Step 3: Compare scenarios

```{r scenario-counts}
scenario_counts <- lapply(
  study$scenarios,
  function(dat) data.frame(
    Rows = nrow(dat),
    Subjects = length(unique(dat$Subject)),
    ScenarioGroup = unique(dat$ScenarioGroup)[1]
  )
)

do.call(rbind, scenario_counts)
```

Scenario data sets contain audit columns when contamination has been applied.
Those audit columns are useful for synthetic diagnostics because ground truth
is known by construction.

```{r audit-columns}
names(study$scenarios$SC2_TimingError)
table(study$scenarios$SC2_TimingError$Contam_Class)
```

## Step 4: Use lower-level pieces when needed

The high-level workflow is a convenience layer. The lower-level generators
remain available when a simulation needs custom templates or custom noise.

```{r low-level}
psi <- simulate_nmea_parameters(n = 8, alpha = 1.2, seed = 202)
observed <- simulate_observed_data(
  psi,
  time_templates = list(
    standard = c(0, 0.5, 1, 3, 6, 9, 12, 15),
    sparse = c(0, 0.75, 3, 8, 14)
  ),
  sigma = function(time) 0.6 + 0.15 * sqrt(pmax(time, 0)),
  seed = 203
)

sparse <- apply_missing_scenario(observed, "S4_SparseMorningMissing", seed = 204)
contaminated <- apply_contamination_scenario(sparse, "SC2_TimingError", seed = 205)
head(contaminated)
```
