---
title: "Synthetic Simulation Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Synthetic Simulation Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
```

```{r setup}
library(cortisol.nmea)
```

This vignette shows the public, synthetic-only simulation workflow. It mirrors
the structure of the Simulation V3 research scripts without depending on real
participant data.

## Step 1: Generate latent subjects

```{r latent}
psi <- simulate_nmea_parameters(n = 40, alpha = 1.2, seed = 101)
latent <- simulate_latent_curves(psi, time_grid = seq(0, 18, by = 0.25))

dim(latent$curve_mat)
```

```{r latent-plot}
matplot(
  latent$time_grid,
  t(latent$curve_mat[1:10, ]),
  type = "l",
  lty = 1,
  col = "#1f6fb255",
  xlab = "Hours since waking",
  ylab = "Cortisol",
  main = "First 10 synthetic latent curves"
)
```

## Step 2: Generate observed data

Observed data are created by sampling time templates and adding measurement
noise. The example below uses a time-varying noise scale.

```{r observed}
time_templates <- list(
  standard = c(0, 0.5, 1, 3, 6, 9, 12, 15),
  early_dense = c(0, 0.33, 0.75, 1.5, 4, 8, 12, 16),
  sparse = c(0, 0.75, 3, 8, 14)
)

sigma_fun <- function(time) {
  0.6 + 0.15 * sqrt(pmax(time, 0)) + 0.35 * exp(-((time - 1) / 0.8)^2)
}

observed <- simulate_observed_data(
  psi,
  time_templates = time_templates,
  sigma = sigma_fun,
  seed = 102
)

head(observed)
```

## Step 3: Apply missingness and sparsity

```{r missing}
sparse <- apply_missing_scenario(
  observed,
  scenario = "S4_SparseMorningMissing",
  seed = 103
)

table(table(sparse$Subject))
```

## Step 4: Apply contamination

```{r contamination}
contaminated <- apply_contamination_scenario(
  sparse,
  scenario = "SC2_TimingError",
  seed = 104
)

table(contaminated$Contam_Class)
```

The contamination generator keeps audit columns, including the clean time and
clean value. These columns are useful for simulation diagnostics and should not
be available in real-data analysis unless ground truth is known by design.

```{r audit-columns}
names(contaminated)
```
