---
title: "Simulation studies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation studies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 3.2)
```

```{r setup}
library(cortisol.nmea)
```

A simulation study asks how well a method recovers curves we know. The package
generates cohorts that resemble prenatal cortisol data, contaminates them,
fits each method and compares the fitted curves with the truth.

## Generating a cohort

`sim_reference()` holds the summaries that drive the generator for one
trimester: the distribution of the curve parameters, the sampling schedule and
its missingness, and the measurement noise. These are aggregate summaries of a
real prenatal cohort; no participant data are included.

```{r reference}
ref <- sim_reference("T2")
ref
```

`sim_cohort()` draws people from it. The observed data and the truth are kept
apart, so the methods only ever see `$observed`.

```{r cohort}
base <- sim_cohort(ref, n = 100, seed = 1)
base
base$observed
```

Samples follow a protocol of waking, +45 min, +2.5 h, +8 h and +12 h, with
realistic delays and about one in five people missing a sample. Noise is larger
in the afternoon than at waking, as in the real data. Any of this can be changed
with `sim_reference_modify()`, for example a noisier assay:

```{r modify}
noisy <- sim_reference_modify(ref, noise_multiplier = 1.5)
```

## Contamination

`sim_contaminate()` adds aberrant samples to a clean cohort. The paper uses
three scenarios: none (D0), 5% of samples raised by 4-8 noise SDs (D1), and 10%
raised by 8-12 noise SDs (D2). Scenarios built from the same cohort share its
people, times and noise, so methods can be compared in pairs.

```{r contaminate}
d2 <- sim_contaminate(base, "D2")
d2
```

`contamination()` defines other patterns: downward or two-sided spikes, spikes
of a fixed size, spikes restricted to a time window, or misrecorded sampling
times.

```{r custom}
contamination(rate = 0.05, magnitude = c(5, 10), scale = "absolute", time_window = c(0, 1))
contamination(rate = 0.05, magnitude = c(0.5, 1.5), type = "timing")
```

## Designing and running a study

A design crosses trimesters, base cohorts (one per seed) and scenarios, and
lists the methods to compare. The default methods are Direct NMEA (no
cleaning), Full NMEA (the full workflow) and a generalized additive mixed model
(GAMM) adapted from Sánchez et al. (2012).

```{r design}
design <- sim_design(trimesters = "T2", n = 40, seeds = list(T2 = 11:13),
                     scenarios = c("D0", "D2"), times = seq(0, 18, by = 0.25))
design
```

```{r run, eval = FALSE}
runs <- sim_run(design, cache_dir = "sim-cache")
```

`sim_run()` saves each data set's results to `cache_dir` as it finishes, so an
interrupted study picks up where it stopped. With `parallel = TRUE` and a
`future::plan()`, base cohorts run in parallel. Failed fits are recorded as
failures and never replaced.

The study above takes a few minutes. Its results ship with the package:

```{r load}
runs <- readRDS(system.file("extdata", "vignette-runs.rds", package = "cortisol.nmea"))
runs
```

## Evaluating

Curves are compared with the truth on a time grid. Within each cohort, errors
are averaged over the people every method could fit, then averaged over
cohorts with a Monte Carlo interval. With only three cohorts here the intervals
are wide; the paper uses ten per trimester.

```{r summary}
s <- eval_summary(runs)$summary
s[s$measure == "rmse", c("scenario", "method", "valid", "mean", "lower", "upper")]
```

`eval_paired()` compares two methods cohort by cohort.

```{r paired}
p <- eval_paired(runs, "Full", "Direct")$summary
p[, c("scenario", "valid", "mean", "lower", "upper", "wins")]
```

Lower errors for the full workflow come partly from excluding people it cannot
fit. Always look at retention alongside recovery.

```{r retention}
r <- eval_retention(runs, "Full")$summary
r[r$scenario == "D2", c("measure", "mean")]
```

`eval_detection()` reports how well the outlier step separated contaminated
from clean samples.

```{r detection}
det <- eval_detection(runs, "Full")$summary
det[det$scenario == "D2" & det$measure %in% c("FPR", "FNR", "roc_auc"),
    c("measure", "mean", "lower", "upper")]
```

## Figures and report

```{r cohort-plot}
plot_cohort_curves(runs, trimester = "T2", seed = 11)
```

```{r subject-plot, fig.width = 6, fig.height = 3.8}
plot_subject(runs, "T2_S0019", trimester = "T2", seed = 11, scenario = "D2")
```

```{r recovery-plot}
plot_recovery(runs)
```

`plot_detection()` and `plot_retention()` show the remaining panels of the
paper's figures. `sim_report(runs, "report.html")` writes all tables and
figures to one HTML file.
