---
title: "Choosing the outlier threshold"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choosing the outlier threshold}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 3.2)
```

```{r setup}
library(cortisol.nmea)
```

The workflow removes a sample when its residual from the initial fit is larger
than a cutoff. A low cutoff removes more contamination but also more good
samples; with only four or five samples per person, each wrongly removed sample
costs a lot, and a person left with fewer than three is excluded altogether.

## The rule

The cutoff is $k$ times a noise standard deviation, so it scales with how noisy
the data are:

```{r rules}
outlier_sd(k = 3.5, sd = "reference")   # trimester reference SD
outlier_sd(k = 3, sd = "iterative")     # SD estimated from the data
outlier_fixed(6)                        # the original fixed cutoff, in nmol/L
```

The default, `outlier_default()`, uses the reference SD for data labelled T1, T2
or T3 and the iterative estimate otherwise. The iterative estimate first sets
aside samples more than 2.75 median absolute deviations from the initial fit,
refits, and takes the residual SD of that fit. It is meant as a first pass that
removes clearly aberrant samples.

Both multipliers were chosen so that the cutoff for second-trimester data is
close to 6 nmol/L, the cutoff of the original analysis. With the reference SD,
the cutoffs are 4.7, 6.2 and 6.8 nmol/L in the first, second and third
trimesters, matching the noise level of each.

## Why not the residual SD of the initial fit

Two other scales are available, `sd = "model"` (the residual SD of the initial
fit) and `sd = "mad"` (the median absolute deviation of its residuals). Both are
estimated from the contaminated fit, so the cutoff rises as contamination rises:
in our development simulation, `"model"` missed 43% of strong spikes, against
25% for the reference SD. `"mad"` removed up to 6% of samples from clean data.
Neither is used by default, but both can be chosen.

## Exploring trade-offs

`tune_threshold()` re-applies different multipliers to the saved initial
residuals of a study. Nothing is refitted, so it is fast, but it reports only
detection errors, not curve recovery.

```{r tune}
runs <- readRDS(system.file("extdata", "vignette-runs.rds", package = "cortisol.nmea"))
tn <- tune_threshold(runs, k = seq(2, 6, by = 0.5), sd = "reference")
d2 <- tn$summary[tn$summary$scenario == "D2", ]
round(d2[, c("k", "cutoff", "FPR", "FNR")], 3)
```

```{r tune-plot}
plot(d2$FPR, d2$FNR, type = "b", pch = 19, xlab = "False positive rate (clean samples removed)",
     ylab = "False negative rate (spikes kept)")
text(d2$FPR, d2$FNR, d2$k, pos = 4, cex = 0.8)
```

The function does not pick a threshold. To compare thresholds by curve recovery
and retention, fit each one:

```{r grid, eval = FALSE}
design <- sim_design(trimesters = c("T1", "T2", "T3"), n = 100, seeds = 1:10,
                     scenarios = c("D0", "D2"),
                     methods = methods_threshold_grid(c(3, 3.5, 4), sd = "reference"))
runs <- sim_run(design, cache_dir = "threshold-cache")
eval_summary(runs)$summary
```

Choose the comparison criteria before looking at the results, and report
retention with recovery.
